\section{Introduction}
\label{sec:intro}
Over the past year, we have heard reports of a number of cases of academic
dishonesty occurring at Worcester Polytechnic Institute. Specifically, most of
these have concerned the unauthorized and unacknowledged copying of source code
within the Electrical and Computer Engineering and Computer Science departments.
The first reports came from Nicholas DeMarinis, a Teaching Assistant in the
Electrical and Computer Engineering department, who noticed several instances
of source code which he considered to be suspiciously similar in an embedded
systems programming course. He identified several cases of academic dishonesty,
including one where the students had extensively obfuscated the copied code.

Meanwhile, Professor Hugh C. Lauer of the Computer Science department found
several instances of academic dishonesty in his own courses. A member of the
course staff of a low-level programming class he was teaching found a pair of
students who submitted near-identical assignments, which he brought to
Professor Lauer's attention. Several assignments later, another member of the
course staff identified another set of students with very similar assignments
--- this time, by noticing that both students had submitted assignments which
had identical, incorrect output.

In both of these cases, the identification of academic dishonesty largely
occurred due to luck. Both courses had more than one teaching assistant; if
the two assignments which contained unauthorized copying were graded by
different teaching assistants, the copying would almost certainly not have been
identified. This draws questions to how many cases of academic dishonesty
occurred which were not identified.

The issue does not appear to be that course staff cannot identify unauthorized
copying when it is present. Instead, there are too many assignments for the
staff to review all of them to identify such cases. The problem is one of data
overload, which is a field computers are well-equipped to deal with. A solution
for flagging potential cases of academic dishonesty for review by course staff
has the potential to greatly increase detection of academic dishonesty.

Given this, Professor Lauer commissioned the authors to construct a system for
automatically detecting suspicious assignments, with the intent of integrating
it into his low-level courses as a tool to identify which assignments required
manual review.


\subsubsection{What is Academic Dishonesty}
\label{sec:lauerdishonesty}
The authors felt that, as they were developing software to detect academic
dishonesty, it was important to know what academic dishonesty is. Professor
Lauer provided us with his definition of what is and is not permissable, and
we used this definition when determining what we should and should not detect.
A wider discussion of varying definitions is contained in \ref{sec:dishonesty}
in the literature review, but this is the working definition we used when
building our solution.

Professor Lauer considers the collaboration of students on a whiteboard to be
perfectly acceptable during assignments (indeed, it may be actively encouraged,
depending on the course). Students can collaboratively develop pseudocode-level
descriptions of the appropriate solution to an assignment, and then translate
this pseudocode into a working program independently. The direct copying of
source code is, however, prohibited --- each student must write their program
independently, though again they may reference a shared pseudocode solution. In
Professor Lauer's view, this ensures that the code must 
``pass through the brain'' --- the students will gain more understanding by
typing independently than by copying code directly. Thus, the direct copying of
code from other sources is considered by Professor Lauer to be academic
dishonesty, while solutions being identical at the pseudocode level are
generally not considered problems.

Using source code from the web follows much the same rules as collaborating
with other students. The use of preexisting pseudocode (or using existing code
as pseudocode) is considered acceptable, so long as the student types their own
solution, and do not copy wholesale a solution from the internet.


\subsection{Project Goals}
\label{sec:goals}
The overall goal of this project was to produce a piece of software to provide
easy flagging of unusually similar source code submissions. The determination of
whether such cases are or are not academic dishonesty is beyond the scope of
this work; it is sufficient that our solution flag them for review by course
staff. A key goal is ease of use --- it must be easy both to use the detector,
and to identify suspicious results it has flagged. To enable features to be
added in the future, we decided that our solution must be open-source and
modular.
% TODO anything else that needs to be said here?
