\documentclass{article}
\author{Dolan Murvihill \and Matthew Heon}
\title{Program Similarity Detection: A review of the literature}
\usepackage{biblatex}
\addbibresource{bibliography.bib}
\newcommand{\vocab}[1]{
  \textit{#1}}
\begin{document}
\maketitle

Detecting similarities between computer programs and source code is a
well-established problem with many applications.

\section{Applications}
\subsection{Code Clone Detection}
Duplicated source code is considered harmful by broad consensus among the
software community, for a number of reasons:
\begin{itemize}
\item Any further changes, especially bug fixes, must me made in both clones;
      often they are
\item [TODO complete this list]
\end{itemize}

Nevertheless, software developers often duplicate code in order to speed up
development, perform experimental changes without impacting existing code, or
to allow two branches to evolve independently \cite{baker95}. Sometimes these
changes are unintentionally left in the source code and cause trouble later.
Software developers would like to reduce the amount of duplicate code in their
software in order to reduce maintenance costs, and the field of \vocab{code
clone detection} has arisen to help locate duplication.

One of the earliest code clone detectors is Dup. Dup's goal is to find code
that is "substantially the same line by line except for global substitutions"
\cite{baker95}. Dup's results were promising for 1995: When run on the source
code for the X Window System, it located 2487 matches of length 30 or more, and
expected that 12\% of the current source code could be eliminated by rewriting
\cite{baker95}. In another system, it found that "two whole directories of 800
lines were \ldots the same except for a systematic change of parameter names
and a line break"\cite{baker95}.

Dup's approach is to locate "maximal exact or parameterized matches" greater
than a threshold length. A \vocab{maximal exact match} means both sections of
text are character for character the same, but the previous and next lines do
not match. \vocab{Parameterized matches}, or P-matches, occur when a one-to-one
mapping of parameter names can transform one section into the other. Dup's
P-matching algorithm is language-dependent and must be modified to support
different types of identifiers.

Dup also includes some other rules to account for oddities like the fact that
even non-duplicate case statements look very similar to one another, or that
a closing brace at the start of a match usually does not belong to duplicated
code.

[detailed description of p-matching detection]

\subsection{Malware Analysis}
\subsection{Dead Code Detection}
\section{Plagiarism Detection}
Plagiarism is the use of another person's work without proper attribution. In
academia, plagiarism is considered a serious integrity violation that is often
penalized with the failure of an assignment or course, and sometimes even
expulsion. Especially in introductory courses, course staff must spend a
significant amount of time looking for relatively unsophisticated plagiarism.
Similarity detection can reduce the amount of time spent on checking for
plagiarism by focusing the course staff on more similar programs without
requiring a pass through the others. Automated similarity measurements can also
improve detection rates by finding similarities that may not be obvious to the
naked eye.

Definitions of plagiarism are many and varied.

% TODO list of definitions
Joy99: "Unacknowledged copying of documents or programs"
Lauer: If students copy the same code from a board, it is kosher
% TODO didn't we find a big UK paper trying to find a unified definition?

%  Reasons for plagiarism
%    Joy99
%      "A weak student produces work in close collaboration with a colleague, in
%        the belief that it is acceptable."
%      "A weak student copies, and then edits, a colleage’s program, with or
%        without the colleage’s permission, hoping that this will go unnoticed."
%      "A poorly motivated (but not necessarily weak) student copies, and then
%        edits, a colleague’s program, with the intention of minimizing the work
%        needed."

The plagiarism detection field contains a number of papers focused on detecting
plagiarism in natural language. Much natural language plagiarism detection
research can be used for source code plagiarism as well, but we tried to focus
on papers that were oriented toward source code.

\subsection{Human analysis requirement}
The proliferation of plagiarism definitions creates a vague, confusing, and
potentially dangerous situation for students. While mitigating the problems
caused by this quagmire is outside the scope of this project, it is important
to recognize that even a detector that is "perfect" (no false positives or
false negatives) under one definition of plagiarism will not be perfect for a
professor using a different definition. What this means is that in making a
final determination on a suspicious pair, there is no substitute for a human.

\subsection{High false positive tolerance}
There is a silver lining, though: because a human is making the final
determination, the cost of a false positive drops from a false accusation of
plagiarism (which is huge) to a small amount of time for the course staff to
review the match. The low cost of a false positive means the checker can use
very sensitive algorithms and be relatively confident that it can catch
(almost) all plagiarism.

\subsection{Obfuscation Techniques}
Plagiarists often try to evade detection by obfuscating their copied document.
In order to assist in developing countermeasures against obfuscation, some
researchers have attempted to characterize these techniques. The best paper on
this topic, by Parker, identifies six "levels" of obfuscation, from a straight,
un-altered file copy (no obfuscation) to sophisticated attacks like refactoring
while loops to for loops and other algebraic translations \cite{parker89}. Any decent
plagiarism detector should be able to counter the lower level normalizations,
but challenges like algebraic refactoring are more difficult for the
implementor as well as the plagiarist.

\section{Known Approaches}
\subsection{Vector distance algorithms}
Vector distance algorithms take vectors of document characteristics and measure
the distance between them.
\subsection{Document fingerprinting}
Document fingerprinting approaches require storing specific elements of one
document or source file, and searching for those elements in another file. One
advantage of most document fingerprinting approaches is that the fingerprints
can be stored in one place, requiring each document to be checked against one
database ($O(n)$ time) rather than comparing each pair of documents ($O(n^2)$).

\subsubsection{COpy Protection System (COPS)}
A good example of a very old document fingerprinting system is COPS. COPS
facilitates \vocab{ordinary operational tests}, including subset, related, and
plagiarism. The system document into \vocab{units}, sequences of words
separated by end-of-sentence markers, and then groups them into, \vocab{chunks}
according to a \vocab{chunking strategy}. Hashes of the chunks were then stored
and retrieved in the database. In COPS, chunks are not related to the
underlying structure of the document.

\subsubsection{n-gram fingerprinting}
By far the most common document fingerprinting approach is called n-gram
fingerprinting. The document is tokenized, and considered in sequences of $n$
tokens, or \vocab{n-grams}. The n-grams overlap, so that there an $m$ character
document will be considered as $m-n+1$ n-grams. The hashes of these n-grams are
stored in a database and checked against those from other documents. The hashes
are usually stored with a pointer to their occurrence to aid in detection.

[illustration here]

Most plagiarism corpora are far too large to store all n-grams from all
documents. Usually, a selection algorithm is used to select a subset of the
n-grams to serve as the document's \vocab{fingerprints}, but that entails
information loss: even word-for-word identical documents will only match if
they both contain an n-gram that was selected as a fingerprint, so how the
fingerprints are selected is very important. A very simple approach is to
select n-grams whose hashes are divisible by some number $p$
\cite{schleimer03}, but this technique can leave gaps of arbitrary length
between fingerprints, so that a large chunk of plagiarized text might slip by.

Schleimer, Wilkerson, and Aiken presented an alternative fingerprint selection
algorithm in 2003 called \vocab{winnowing}. After dividing the document's
tokens into n-grams, winnowing further divides the n-grams into \vocab{windows}
of size $t$. Winnowing always selects the minimum hash from each window,
guaranteeing that a match of $t$ or more consecutive tokens will be identified.

[illustration here]

Winnowing has proven to be extremely powerful; it is used by the current
industry standard source code similarity detector, MOSS, where the authors
claim that "false positives appear to be non-existent, and the infrequent
reports of false negatives have always been tracked back either to
implementation bugs or user error" \cite{schleimer03}. We doubt the picture is
that spectacular, but there is no doubt that MOSS is a very successful
platform.

\subsection{Common substring techniques}
<explanation of Irving '04 paper>
\subsection{Feature Analysis}

%Substring
%    Basis of Work: Irving '04 paper
%    Tokenization Approaches - Whitespace vs Character
%    Algorithm Complexity
%  Longest Common Substring
%    Want to implement - do we have time?
%    Faster but less accurate than Smith-Waterman

\printbibliography
\end{document}
