\documentclass{article}
\title{Software Similarity Detection: Proposal}
\author{Matthew Heon \and Dolan Murvihill}

\begin{document}
\maketitle
Academic dishonesty is a common phenomenon in any education system. It is most
common in introductory classes, where struggling students duplicate the work of
other, more talented students and claim it as their own. Plagiarism severely
damages the integrity of a class, and is considered one of the most severe
transgressions in academia. The severity of unauthorized copying, coupled with
its frequency, means that instructors and teaching assistants must spend a
significant amount of time checking for academic dishonesty. Even so, cheating
is not always caught.

To help save instructors' time and reduce the frequency
of uncaught duplication, we propose automatically applying heuristics to
identify similar work in the specific discipline of computer programming. No
current technology can replace a human in determining whether a work was
copied, but we believe appropriate algorithms can focus the attention of the
course staff on similar work submissions, allowing them to catch more with the
time they have. This framework will also allow the detection of instances of the
so-called ``copy-and-paste pattern'' where identical code is reused in multiple
places, a useful metric for large projects.
% Added another potential application... Not sure if worth mentioning?

\section{Deliverables}
This is a two-term MQP, with 2/3 credits being applied in A term of 2014 and
1/3 credit being applied in B term of 2014. Over the course of the project,
we aim to produce several deliverable products, summarized below.
% Reworded to avoid dangling end-of-sentence

\subsection{Literature Review}
We will conduct a thorough review of prior work in the field of plagiarism
detection, and the domain-specific field of programming languages, to identify
the most common techniques in use today and inform our direction. We will
produce a summary of our findings, discussing relevant algorithms and
evaluation techniques, their tradeoffs, and any previous implementations of
similar software. % Reword, adding note that we're researching previous impls
Our findings will be thoroughly cited, with seminal papers in the field being
mentioned specifically to ease any future search for information on the subject.
The finished literature review will be ready at the end of A term, but we aim
to complete the majority of research early in the term so it may be used to
engineer the software deliverables.
% Reworded, added reason why we're calling our seminal works

\subsection{Evaluation Framework}
We will design a software system which we can use to receive measurable,
reproducible, actionable data to evaluate any algorithms we implement. 
% Unlikely we're designing any algorithms, that's pretty difficult.
% Also, fixed tense issue below.
The framework will be designed to give us a quantitative understanding of our
product's characteristics, which will inform our engineering process. The
framework will be designed to conform as closely as possible to the industry
standard methods for evaluating similar software, as identified in our
literature review. The evaluation framework will be built early in A term to
enable its use in continuous integration testing of the other software
deliverables, and we will improve it continuously as needed.
% Clarify that we're building this first

\subsection{Server Framework}
% Clarify we're just doing Linux support. I am not interested in explicitly
% supporting Solaris and the BSD family, nor do I feel it adds value to our
% proposal.
We will create server software, designed to run on the Linux operating
system, that provides an interface for software similarity checking and a module
system to allow the easy addition of similarity detection algorithms. 
The server will provide an HTTP REST API for requesting similarity
checks on assignments. THe client will be allowed to specify relevant parameters
for the checks - for example, specifying which algorithms should be used, or
overriding default alert thresholds. The server will also be able to store
a large volume of previously-submitted work to check against. We aim to produce
a compact, specialized framework well-suited for this task. Finally, we will
produce documentation to enable the easy extension of the server with new
detection algorithms.
% Didn't know what you were getting at with the compact-as-possible thing,
% so I tried to clarify - is this what you meant?

\subsection{Example Plugin}
We will implement a simple similarity-checking algorithm focused on detecting
simple code duplication. This algorithm will provide a means of testing our
framework and will implement a baseline for detection functionality. At the end
of B term, we intend for the combination of our server and this plugin to be
usable as a means of detecting simple instances of plagiarism. We aim to
complete this plugin by the beginning of B term to enable testing of the
framework on real source code.
% I'm pretty sure this was desired for C term, not B?
% Otherwise, major reword to make it less advisor-focused, more general
% application - we want it to be ready for simple applications everywhere when
% we're done.

\subsection{Example Client}
We will develop a command line tool, available on UNIX and Windows, that makes
appropriate REST calls to the server and reports the responses in a
human-friendly manner. We aim to have this client available at the end of C
term.

\subsection{Final Paper}
We will write a final paper according to WPI's project guidelines,
incorporating our literature review, the process used to build our software
deliverables, and details of the functionality and effectiveness of the final
deliverables.

\end{document}
